The role of advection versus diffusion in intraventricular transport

### 2025-02-03 paper/main.tex
methods: update experimental, switch order experimental-computational

% Comments
\newcommand{\lyng}[1]{\textcolor{blue}{#1}}
\newcommand{\rt}[1]{\textcolor{red}{#1}}
\newcommand{\mer}[1]{\textcolor{magenta}{#1}}
\newcommand{\mk}[1]{\textcolor{olive}{MK: #1}}

@@

% Title
\title{The role of advection versus diffusion in intraventricular transport}
\author[1,*]{Halvor Herlyng}
\author[1]{Ada J.~Ellingsrud}
\author[1]{Miroslav Kuchta} 
\author[X]{Inyoung Jeong} % OK?
\author[1,3,+]{Marie E.~Rognes} % OK order?
\author[2,+]{Nathalie Jurisch-Yaksi}
\affil[1]{Department for Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway}
\affil[2]{Department of Clinical and Molecular Medicine, Norwegian University of Science and Technology, Trondheim, Norway}
\affil[3]{K.~G.~Jebsen Center for Brain Fluid Research}
\affil[*]{\href{mailto:hherlyng@simula.no}{hherlyng@simula.no}}
\affil[+]{these authors contributed equally to this work}
% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={zebrafish-cilia-csf},
  pdfauthor={H.~Herlyng \textit{et al.}}
}
\fi

\begin{abstract}
    This paper presents ...
    %Flow of cerebrospinal fluid in zebrafish brain ventricles is studied in this work. Motivated by experimental observations, we consider motile cilia and cardiac beating as the driving flow mechanisms. A mathematical model of the flow is developed and coupled with a transport model to study the transport of molecules inside the brain ventricles. The Stokes equations are used to model the flow, with the motile cilia acting as a force on the fluid through a stress boundary condition that allows slip at the wall. Transport is modeled with an advection-diffusion equation coupled to the velocity field obtained from solving the Stokes equations. The flow and transport models are discretized with finite element methods. For the Stokes equations we employ a non-conforming discretization based on elements of the Brezzi-Douglas-Marini family. The advection-diffusion equation is discretized with discontinuous Galerkin elements. We implement and solve the discretized equations numerically to simulate flow and transport in a zebrafish brain ventricles geometry obtained with medical imaging.
\end{abstract}
\begin{document}
\maketitle
% Introduction:
% 1. What are cilia and why are they important?
% 2. Previous studies of cilia flow
% 3. What problem do we want to solve
% 4. The model and work done in this paper
\section*{Introduction}
Brain development and function depend on numerous regulatory processes, including interactions with cerebrospinal fluid (CSF).
CSF is a clear liquid produced by the choroid plexus, which is a secretory epithelium located within the brain ventricles.
Following production, CSF flows through the ventricular system and subarachnoid space, where it regulates brain development and
homeostasis by transporting growth factors, nutrients, neurotransmitters, waste products and other molecules and metabolites.

@@

The technical limitations of studying cilia-mediated flow and transport opens the possibility of computational investigations. Cilia have been studied previously with mathematical modeling,
with numerous studies on the interaction between single or several cilia and surrounding fluids~\cite{Guo2020SimulatingGeometries, Ruvalcaba2021NumericalTree, Smith2009MathematicalFluids, Cui2019NumericalMethod, Cui2022AFlow}.
Others consider applications such as ciliary transport of mucous in airways~\cite{Fulford1986Muco-ciliaryLung, Ramirez-SanJuan2020Multi-scaleArrays},
propulsion and locomotion of ciliated organisms~\cite{BLAKE1974MechanicsMotion, Jahn1972LocomotionProtozoa, Brennen1977FluidFlagella},
coral reefs~\cite{Pacherres2022CiliaryProduction}, and the brain~\cite{Siyahhan2014FlowVentricles, Yoshida2022EffectVentricles, Salman2022ComputationalEmbryo, Thouvenin2020OriginCanal}.

@@

\subsection*{Zebrafish maintenance and strains in photoconversion experiments}
The zebrafish~(\emph{danio rerio}) animal facilities used are approved by the Norwegian Food Safety Authority (NFSA, Mattilsynet). Furthermore, the zebrafish were maintained in accordance with the guidelines set by the NFSA and the European Communities Council Directive. The larval and adult zebrafish were raised under standard husbandry conditions at 28.5°C in a Techniplast Zebtech Multilinking system. The fish tanks were kept at constant pH 7.0 and 685 $\mu$S with a 14/10 hr light/dark cycle. From fertilization to 3 dpf, larvae were maintained in egg water (1.2 g marine salt and 0.1\% methylene blue in 20 L reverse osmosis (RO) water) and subsequently transferred to artificial fish water (AFW) (1.2 g marine salt in 20 L RO water). The zebrafish lines used in the experiments were \emph{ccdc103(dnaaf19)\textsuperscript{tn222a}} (\emph{schmalhans (smh)})~\cite{Jau-NianChen1997Left-rightZebrafish}(received from Drummond lab, MGH) and \emph{Tg(gfap:Gal4FF)\textsuperscript{nw7Tg}}~\cite{DiazVerdugo2019Glia-neuronSeizures}, \emph{Tg(5xuas:Signal-Dendra2)\textsuperscript{nw21Tg}}. Animals were in the pigmentless \emph{nacre\textsuperscript{b692}}(\emph{mitfa\textsuperscript{-/-}})~\cite{JamesA.Lister1999NacreFate} background.

Whole-mount~\emph{in vivo} live imaging and photoconversion experiments were performed with zebrafish larvae at the 2 days post-fertilization (dpf) stage obtained from inbreeding of heterozygous \emph{smh\textsuperscript{+/-};Tg(gfap:Gal4FF);Tg(5xuas:Signal-Dendra2}) adult animals, as described by D’Gama~\emph{et al.}~\cite{DGama2024Cilia-mediatedBrain}. Controls consisted of either wild-type (\emph{smh}\textsuperscript{+/+}) or \emph{smh}\textsuperscript{+/-} heterozygous zebrafish from the same breeding. The mutants were identified based on their curved body~\cite{Jau-NianChen1997Left-rightZebrafish}.

\subsection*{Genotyping}
Genomic DNA (gDNA) was isolated from clipped fins of anesthetized adult fish using 100 $\mu$L of PCR lysis buffer (containing 1M tris pH 7--9, 0.5 M EDTA, tritonX-100, and Proteinase K 0.1 mg/ml) overnight at 50°C. To stop the lysis reaction, the samples were heated to 95°C for 10 minutes and then centrifuged at 13000 rpm for 2 minutes. The supernatant containing gDNA was utilized for KASP assays-based analysis. The gDNAs were diluted (1:2) with water, and 3 $\mu$L of the diluted gDNA was used to perform the KASP assay following the guidelines of the manufacturer (LGC Biosearch Technologies™). For each sample well, the master mix contained: 5 $\mu$L of master mix; 0.14 $\mu$L of assay mix; 1.86 $\mu$L of RO water.

\subsection*{Generation of \emph{Tg(5xuas:Signal-Dendra2)\textsuperscript{nw21Tg}} line}
\mer{MER: Could this heading be made understandable by non-experts?}
The open reading frame (ORF) of the neuropeptide y (npy) signal peptide (\emph{npy}: ENSDARG00000036222, Q1LW93) was fused with the N-terminal of zebrafish codon optimized Dendra2 DNA sequence. Next the conjugated DNA sequence was synthesized with EcoR1 enzyme sites at the 5’ and 3’ ends (GenScript Biotech), and the synthesized signal-Dendra2 DNA was inserted into the pT2MUASMCS vector~\cite{Asakawa2008GeneticZebrafish} through restriction enzyme cloning (GenScript Biotech).  To generate the transgenic line, the 5xuas:Signal-Dendra2 plasmid DNA was used. A volume of 2 nl of a mixture of the plasmid DNA (60 pg) and tol2 mRNA (10 pg) was microinjected into one-cell stage embryos, as described by Jeong~\emph{et al.}~\cite{Jeong2024TheZebrafish}. The injected embryos were raised to adulthood (F0). Germline-transmitted founder zebrafish were identified by breeding with multiple Gal4 transgenic lines. Stable F1 embryos expressing the Gal4-driven Dendra2 signals were screened and raised to adult zebrafish.

\subsection*{Whole-mount zebrafish \emph{in vivo} live time-lapse imaging and photoconversion of Dendra2}
To image 2 dpf larval zebrafish, the fish were anesthetized in 0.013\% MS222 in AFW and mounted laterally in 1.5\% low melting agarose in a Flurodish (VWR, FD35PDL-100), to obtain a lateral view of the brain ventricles. After the fish were positioned in the agarose, the dish stood for 5 min to allow the agarose to solidify. After solidification, AFW containing 0.013\% MS222 was added, and then the fish were transferred to the confocal microscope (LSM880 Examiner, Zeiss) and imaged using a 20x water-immersion objective (Zeiss, NA 1.0, Plan-Apochromat) at room temperature.

The time-lapse images were acquired in a single plane, covering the telencephalic, diencephalic and rhombencephalic ventricles simultaneously, with a frequency of 0.37--0.38 Hz (2.60--2.68 sec / frame). The size of the images was $1024\times400$ pixels. In total, 300 images were acquired per fish. The first 10 images were scanned without photoconversion to obtain a baseline value of fluorescence intensity, and the next 290 images were scanned while performing photoconversion. For photoconversion, ‘Bleaching’ and ‘Region’ in the ZEN software were used. A 405 nm laser was focused into a circular area (diameter: 16 $\mu$m, scan 0.77 $\mu$sec/pixel) of the dorsoposterior diencephalic ventricle with 100\% laser power. The laser was illuminated in the area repeatedly between each scanning. After imaging, the fish health was checked and only data from healthy fish were analyzed.
Data were collected and analyzed from three separate experiments, with a total of 8 controls and 6 mutants.

\subsection*{Photoconversion data analysis}

Photoconversion data analysis was performed in MATLAB and the codes are \fixme{openly available\cite{zenodo-archive-for-repo-link-fixme}}. Following import of the image data, the photoconverted channel was aligned to correct for drift in $x,y$ directions using previously developed codes~\cite{Reiten2017Motile-Cilia-MediatedComputations, Ringers2023NovelEpithelia}. Only stable recordings were further analyzed. To increase the signal-to-noise ratio, the images were downsampled by a block average using a resampling factor of 6 (using the MATLAB function blockproc). The change of fluorescence $($$(F-F_0)/F_0$, hereafter referred to as $\Delta F$$)$ was calculated for each pixel at each time point of the data. Here, the value $F_0$ is the average fluorescence intensity for the baseline (the first 10 time-lapse images) acquired before photoconversion.

In order to obtain $\Delta F$ values for a region of interest (ROI), ROIs were first drawn on the aligned and block-averaged time series. Six ROIs were drawn manually: ROI 1 at the location of photoconversion, and ROIs 2--6 in different regions of the ventricular system. The pixel values of the fluorescence intensity within one ROI are averaged for each time point, and the relative change $\Delta F$ is then calculated with this averaged value. To normalize the data with respect to photoconversion efficiency, the $\Delta F$ curves for all ROIs were divided by the $\Delta F$ values obtained for ROI 1 at the end of the experiment, so that $\Delta F$ values for the photoconverted site ROI 1 ranged from zero to one. Finally, we average the $\Delta F$ data from controls over the number of controls, and the data from mutants over the number of mutants. We report these mean values together with a one-standard deviation error band.

\subsection*{Intraventricular injection of microbeads and particle tracking}
In addition to photoconversion experiments, we carried out particle tracking inside the brain ventricles of a larval zebrafish, to demonstrate the motile-cilia and cardiac-cycle related flow features observed and described in Olstad~\emph{et al.}~\cite{Olstad2019CiliaryDevelopment}. A schematical illustration of the flow features in the brain ventricles is shown in~\Cref{fig:figure1}a. Our experimental setup for the particle tracking is identical to the ones considered there (\lyng{HH: Is this true?}), and we refer to that paper for further details than the ones provided here in the following.

Anesthetized 2 dpf control zebrafish were injected with 1 nl of a mixture containing 0.1\% w/v fluorescent beads (SPHERO Fluorescent Yellow Particles 1\% w/v, $F = 0.16$ mm) diluted in 7.5 mg/ml 70 kDa rhodamine B isothiocyanate-dextran (RITC-dextran; Sigma-Aldrich, R9379) dissolved in artificial CSF. Artificial CSF composition was as follows: 124 mM NaCl, 22 mM D-(+)-Glucose, 2.0 mM KCl, 1.6 mM MgSO\textsubscript{4} $\cdot$ 7 H\textsubscript{2}O, 1.3 mM KH\textsubscript{2}PO\textsubscript{4}, 24 mM NaHCO\textsubscript{3}, 2.0 mM CaCl\textsubscript{2} $\cdot$ 2 H\textsubscript{2}O. The needles used for the injections were pulled with a Sutter Instrument Co. Model P-2000 from thin-walled glass capillaries (1.00 mm; VWR) and cut open with a forceps. A volume of 1 nl of the solution was injected with a pressure injector (Eppendorf Femtojet 4i) in the rostral rhombencephalic ventricle. The pressure and time were calibrated for each needle using a 0.01 mm calibration slide.

Following injection, the zebrafish was directly placed under the confocal microscope and 1200 images were acquired at a frequency of 13.1 Hz. A single optical section was obtained. Particle tracking was done using TrackMate~\cite{Tinevez2017TrackMate:Tracking} in Fiji/ImageJ~\cite{Schindelin2012Fiji:Analysis} and plotted using MATLAB, as described by Jeong~\emph{et al.}~\cite{Jeong2022MeasurementTelencephalon}. The parameters for TrackMate were as follows. The DoG detector was used with the settings: threshold: 2.0, with median filtering; radius: 2.0, with subpixel localization. Next, the Simple LAP tracker was used with settings: max frame gap: 2; alternative linking cost factor: 1.05; linking max distance: 5.0; gap closing max distance: 2.0; splitting max distance: 15.0; allow gap closing: true; allow track splitting: false; allow track merging: false; merging max distance: 15.0; cutoff percentile: 0.9. Only the particle tracks with at least 30 data points, while simultaneously covering a distance of at least 8.8 $\mu$m, were shown (\cref{fig:figure1}, panels \textbf{b} and \textbf{c}).

@@

To computationally represent protein photoconversion in a region $\Omega_{\mathrm{pc}} \subset \Omega$ (corresponding to ROI 1), we prescribe a given time-dependent value to the solute concentration in this region. 
% in $\Omega_{\mathrm{pc}}$.
%, such that $c(\xx, t)$ for all $\xx\in\Omega_{\mathrm{pc}}$ and $t>0$.
Mimicking the experimental protocol, we consider the dorsal part of the diencephalic ventricle as the subdomain $\Omega_{\mathrm{pc}}$, and set
\begin{equation}
  c(\xx, t)=\frac{\log{(1+t/a)}}{\log{(1+T/a)}}
  \qquad \xx \in \Omega_{\mathrm{pc}}, \quad \fixme{t > 0}.
\label{eq:injection_curve}
\end{equation}
The value $a=65$ was chosen to fit the fluorescence intensity curve observed in the physical photoconversion experiments.

@@

>>>>

% Comments
\newcommand{\lyng}[1]{\textcolor{blue}{#1}}
\newcommand{\mer}[1]{\textcolor{magenta}{#1}}
\newcommand{\mk}[1]{\textcolor{olive}{MK: #1}}

@@

% Title
\title{The role of advection versus diffusion in intraventricular transport}
\author[1,*]{Halvor Herlyng}
\author[1]{Ada J.~Ellingsrud}
\author[1]{Miroslav Kuchta} 
\author[X]{Inyoung Jeong} % OK?
\author[1,3,+]{Marie E.~Rognes} % OK order?
\author[2,+, *]{Nathalie Jurisch-Yaksi}
\affil[1]{Department for Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway}
\affil[2]{Department of Clinical and Molecular Medicine, Norwegian University of Science and Technology, Trondheim, Norway}
\affil[3]{K.~G.~Jebsen Center for Brain Fluid Research}
\affil[*]{\href{mailto:hherlyng@simula.no}{hherlyng@simula.no} \href{mailto:nathalie.jurisch-yaksi@ntnu.no}{nathalie.jurisch-yaksi@ntnu.no}}
\affil[+]{these authors contributed equally to this work}
% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={zebrafish-cilia-csf},
  pdfauthor={H.~Herlyng~\emph{et al.}}
}
\fi

\begin{abstract}
    This paper presents ...
    %Flow of cerebrospinal fluid in zebrafish brain ventricles is studied in this work. Motivated by experimental observations, we consider motile cilia and cardiac beating as the driving flow mechanisms.
    %A mathematical model of the flow is developed and coupled with a transport model to study the transport of molecules inside the brain ventricles. 
    %The Stokes equations are used to model the flow, with the motile cilia acting as a force on the fluid through a stress boundary condition that allows slip at the wall. 
    %Transport is modeled with an advection-diffusion equation coupled to the velocity field obtained from solving the Stokes equations.
    %The flow and transport models are discretized with finite element methods. For the Stokes equations we employ a non-conforming discretization based on elements of the Brezzi-Douglas-Marini family. 
    %The advection-diffusion equation is discretized with discontinuous Galerkin elements. 
    %We implement and solve the discretized equations numerically to simulate flow and transport in a zebrafish brain ventricles geometry obtained with medical imaging.
\end{abstract}
\begin{document}
\maketitle
% Introduction:
% 1. What are cilia and why are they important?
% 2. Previous studies of cilia flow
% 3. What problem do we want to solve
% 4. The model and work done in this paper
\section*{Introduction}
Brain development and function depend on numerous regulatory processes, including interactions with cerebrospinal fluid (CSF).
CSF is a clear liquid produced by the choroid plexus, which is a secretory epithelium located within the brain ventricles.
Following production, CSF flows through the ventricular system and subarachnoid space, where it regulates brain development and
homeostasis by transporting growth factors, nutrients, neurotransmitters, waste products and other molecules and metabolites.

@@

The technical limitations of studying cilia-mediated flow and transport opens the possibility of mathematical modeling and computational investigations. Cilia have been studied previously,
with numerous studies on the interaction between single or several cilia and surrounding fluids~\cite{Guo2020SimulatingGeometries, Ruvalcaba2021NumericalTree, Smith2009MathematicalFluids, Cui2019NumericalMethod, Cui2022AFlow}.
Others considered applications such as ciliary transport of mucous in airways~\cite{Fulford1986Muco-ciliaryLung, Ramirez-SanJuan2020Multi-scaleArrays},
propulsion and locomotion of ciliated organisms~\cite{BLAKE1974MechanicsMotion, Jahn1972LocomotionProtozoa, Brennen1977FluidFlagella},
coral reefs~\cite{Pacherres2022CiliaryProduction}, and the brain~\cite{Siyahhan2014FlowVentricles, Yoshida2022EffectVentricles, Salman2022ComputationalEmbryo, Thouvenin2020OriginCanal}.
These studies have provided valuable insights, showing that cilia lend themselves to simulation studies.

@@

@@

To computationally represent protein photoconversion in a region $\Omega_{\mathrm{pc}} \subset \Omega$ (corresponding to ROI 1), we prescribe a given time-dependent value to the solute concentration in this region. 
Mimicking the experimental protocol, we consider the dorsal part of the diencephalic ventricle as the subdomain $\Omega_{\mathrm{pc}}$, and set
\begin{equation}
  c(\xx, t)=\frac{\log{(1+t/a)}}{\log{(1+T/a)}}
  \qquad \xx \in \Omega_{\mathrm{pc}}, \quad \fixme{t > 0}.
\label{eq:injection_curve}
\end{equation}
The value $a=65$ was chosen to fit the fluorescence intensity curve observed in the physical photoconversion experiments.

@@

%%------PHYSICAL EXPERIMENTS------%%
\subsection*{Zebrafish maintenance and strains in photoconversion experiments}
The animal facilities for zebrafish~(\emph{danio rerio}) are approved by the Norwegian Food Safety Authority (NFSA, Mattilsynet).
The zebrafish were maintained in accordance with the guidelines set by the NFSA and the European Communities Council Directive.

The larval and adult zebrafish were raised under standard husbandry conditions at 28.5°C in a Techniplast Zebtech Multilinking system.
The fish tanks were kept at constant pH 7.0 and 685 $\mu$S with a 14/10 hr light/dark cycle.
From fertilization to 3 dpf, larvae were maintained in egg water (1.2 g marine salt and 0.1\% methylene blue in 20 L reverse osmosis (RO) water)
and subsequently transferred to artificial fish water (AFW) (1.2 g marine salt in 20 L RO water).
The zebrafish lines used in the experiments were \emph{ccdc103(dnaaf19)\textsuperscript{tn222a}} (\emph{schmalhans (smh)})~\cite{Jau-NianChen1997Left-rightZebrafish}
and \emph{Tg(gfap:Gal4FF)\textsuperscript{nw7Tg}}~\cite{DiazVerdugo2019Glia-neuronSeizures}, \emph{Tg(5xuas:Signal-Dendra2)\textsuperscript{nw21Tg}}.
Animals were in the pigmentless \emph{nacre\textsuperscript{b692}}(\emph{mitfa\textsuperscript{-/-}})~\cite{JamesA.Lister1999NacreFate} background.

Wholemount~\emph{in vivo} live imaging and photoconversion experiments were performed with zebrafish larvae at 2 days post-fertilization (dpf) stage obtained from inbreeding of heterozygous
\emph{smh\textsuperscript{+/-};Tg(gfap:Gal4FF);Tg(5xuas:Signal-Dendra2}) adult animals, as described in D'Gama~\emph{et al.}~\cite{DGama2024Cilia-mediatedBrain}.
Controls consisted of either wild-type (\emph{smh}\textsuperscript{+/+}) or \emph{smh}\textsuperscript{+/-} heterozygous zebrafish from the same breeding.
The mutants were identified based on their curved body~\cite{Jau-NianChen1997Left-rightZebrafish}.

\subsection*{Genotyping of adult zebrafish}
Genomic DNA (gDNA) was isolated from clipped fins of anesthetized adult fish using 100 $\mu$L of PCR lysis buffer (containing 1M tris pH 7--9, 0.5 M EDTA, tritonX-100, and Proteinase K 0.1 mg/ml) overnight at 50°C.
To stop the lysis reaction, the samples were heated to 95°C for 10 minutes and then centrifuged at 13000 rpm for 2 minutes.
The supernatant containing gDNA was utilized for KASP assays-based analysis.
The gDNAs were diluted (1:2) with water, and 3 $\mu$L of the diluted gDNA was used to perform the KASP assay following the guidelines of the manufacturer (LGC Biosearch Technologies™).
For each sample well, the master mix contained: 5 $\mu$L of master mix; 0.14 $\mu$L of assay mix; 1.86 $\mu$L of RO water.

\subsection*{Generation of \emph{Tg(5xuas:Signal-Dendra2)\textsuperscript{nw21Tg}} line}
\lyng{Needs a more understandable title.}
The open reading frame (ORF) of the neuropeptide y (npy) signal peptide (\emph{npy}: ENSDARG00000036222, Q1LW93) was fused with the N-terminal of zebrafish codon optimized Dendra2 DNA sequence.
The DNA sequence was synthesized with EcoR1 enzyme sites at the 5’ and 3’ ends (GenScript Biotech),
and inserted into the pT2MUASMCS vector~\cite{Asakawa2008GeneticZebrafish} through restriction enzyme cloning (GenScript Biotech).

To generate the transgenic line, the 5xuas:Signal-Dendra2 plasmid DNA was used.
A volume of 2 nl of a mixture of the plasmid DNA (60 pg) and tol2 mRNA (10 pg) was microinjected into one-cell stage embryos,
as described in Jeong~\emph{et al.}~\cite{Jeong2024TheZebrafish}. The injected embryos were raised to adulthood (F0).
Germline-transmitted founder zebrafish were identified by breeding with multiple Gal4 transgenic lines. Stable F1 embryos expressing the Gal4-driven Dendra2 signals were screened and raised to adult zebrafish.

\subsection*{Wholemount zebrafish \emph{in vivo} live time-lapse imaging and photoconversion of Dendra2}
Two dpf larval zebrafish were anesthetized in 0.013\% MS222 (Ethyl 3-aminobenzoate methanesulfonate, Sigma) in AFW and mounted laterally in 1.5\% low melting agarose in a Flurodish (VWR, FD35PDL-100),
to obtain a lateral view of the brain ventricles. After the fish were positioned in the agarose, the dish stood for 5 min to allow the agarose to solidify.
After solidification, AFW containing 0.013\% MS222 was added, and then the fish were transferred to the confocal microscope (LSM880 Examiner, Zeiss)
and imaged using a 20x water-immersion objective (Zeiss, NA 1.0, Plan-Apochromat) at room temperature.

The time-lapse images were acquired in a single plane, covering the telencephalic, diencephalic and rhombencephalic ventricles simultaneously,
with a frequency of 0.37--0.38 Hz (2.60--2.68 sec / frame). The size of the images was $1024\times400$ pixels. In total, 300 images were acquired per fish.
The first 10 images were scanned without photoconversion to obtain a baseline value of fluorescence intensity, and the next 290 images were scanned while performing photoconversion.

For photoconversion, ‘Bleaching’ and ‘Region’ in the ZEN software were used. A 405 nm laser was focused into a circular area (diameter: 16 $\mu$m, scan 0.77 $\mu$sec/pixel)
of the dorsoposterior diencephalic ventricle with 100\% laser power. The laser was illuminated in the area repeatedly between each scanning.
After imaging, the fish health was checked and only data from healthy fish were analyzed.
Data were collected and analyzed from three separate experiments, with a total of 8 controls and 6 mutants.

\subsection*{Photoconversion data analysis}
All data analysis was performed in MATLAB and the codes are available in the repository of this article: \url{repo-link}.

Following import of the image data, the photoconverted channel was aligned to correct for drift
in $x,y$ directions using previously developed codes~\cite{Reiten2017Motile-Cilia-MediatedComputations, Ringers2023NovelEpithelia}. Only stable recordings were further analyzed. 
To increase the signal to noise ratio, the images were downsampled by a block average using a resampling factor of 6 (using the MATLAB function blockproc).
The change of fluorescence $($$(F-F_0)/F_0$, hereafter referred to as $\Delta F$$)$ was calculated for each pixel at each time point of the data.
Here, the value $F_0$ is the average fluorescence intensity for the baseline (the first 10 time-lapse images) acquired before photoconversion.

To obtain $\Delta F$ values for a region of interest (ROI), ROIs were first drawn on the aligned and block-averaged time series.
Six ROIs were drawn manually: ROI 1 around the location of photoconversion, and ROIs 2--6 in different regions of the ventricular system.
The pixel values of the fluorescence intensity within one ROI were averaged for each time point, and the relative change $\Delta F$ calculated with this averaged value.
To normalize the data with respect to photoconversion efficiency, the $\Delta F$ curves for all ROIs were divided by the $\Delta F$ values obtained for ROI 1 at the end of the experiment,
so that $\Delta F$ values for the photoconverted site ROI 1 ranged from zero to one. Finally, we averaged the $\Delta F$ data from controls and from mutants.
We report these mean values together with a one-standard deviation error band.

\subsection*{Intraventricular injection of microbeads and particle tracking}
To recover the flow fields shown in~\Cref{fig:figure1}b, c, we carried out particle tracking in a larval zebrafish injected with fluorescent beads.
As described in Olstad~\emph{et al.}~\cite{Olstad2019CiliaryDevelopment}, anesthetized 2 dpf zebrafish were injected with 1 nl of a mixture containing 0.1\% w/v fluorescent beads
(SPHERO Fluorescent Yellow Particles 1\% w/v, F = 0.16 mm) diluted in 7.5 mg/ml 70 kDa rhodamine B isothiocyanate-dextran (RITCdextran; Sigma-Aldrich, R9379)
dissolved in artificial CSF. Artificial CSF composition was as follows: 124 mM NaCl, 22 mM D-(+)-Glucose, 2.0 mM KCl, 1.6 mM MgSO\textsubscript{4} $\cdot$ 7 H\textsubscript{2}O,
1.3 mM KH\textsubscript{2}PO\textsubscript{4}, 24 mM NaHCO\textsubscript{3}, 2.0 mM CaCl\textsubscript{2} $\cdot$ 2 H\textsubscript{2}O.
The needles used for the injections were pulled with a Sutter Instrument Co. Model P-2000 from thin-walled glass capillaries (1.00 mm; VWR) and cut open with a forceps.
A volume of 1 nl of the solution was injected with a pressure injector (Eppendorf Femtojet 4i) in the rostral rhombencephalic ventricle.
The pressure and time were calibrated for each needle using a 0.01 mm calibration slide.

Following injection, the zebrafish was directly placed under the confocal microscope and 1200 images were acquired at a frequency of 13.1 Hz.
A single optical section was obtained. Particle tracking was done using TrackMate~\cite{Tinevez2017TrackMate:Tracking} in Fiji/ImageJ~\cite{Schindelin2012Fiji:Analysis}
and plotted using MATLAB, as described in Jeong~\emph{et al.}~\cite{Jeong2022MeasurementTelencephalon}. The parameters for TrackMate were as follows.
The DoG detector was used with the settings: threshold: 2.0, with median filtering; radius: 2.0, with subpixel localization.
Next, the Simple LAP tracker was used with settings: max frame gap: 2; alternative linking cost factor: 1.05; linking max distance: 5.0; gap closing max distance: 2.0;
splitting max distance: 15.0; allow gap closing: true; allow track splitting: false; allow track merging: false; merging max distance: 15.0; cutoff percentile: 0.9.
Only the particle tracks with at least 30 data points, while simultaneously covering a distance of at least 8.8 $\mu$m, were plotted in~\Cref{fig:figure1}b, c.
